FraCaS: Temporal Analysis

### 2020-11-09 paper2020/paper.tex
extend discussion on progressive

We treat verbs in the progressive form as different semantically from
the non-progressive form. For example, ``John was writing a book'' is
encoded as
$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, PROG\_write(John,book,t_1,t_2)$,
while ``John wrote a book'' is encoded as
$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, write(John,book,t_1,t_2)$. This is
necessary, because $PROG\_write$ is subject to
$\constant{subsumption}$: if John is writing in the interval
$[t_1,t_2]$ then he is writing in any sub-interval of
$[t_1,t_2]$. However, the same cannot be said of $write$: the
end-points of the interval indicate the time needed to complete the
achievement. (For example, ``John wrote a book in 1993'' neither
entails ``John wrote a book in January 1993'' nor ``John wrote a book
in December 1993''.) In fact, \emph{write}, in the non-progressive
from, is on the contrary subject to $\constant{unicity}$. Finally, we
also have $write(x,y,t_1,t_2) → PROG\_write(x,y,t_1,t_2)$. That is,
the achievement (or \emph{activity} in our terminology) variant implies the stative variant, in the
same interval.  Consequently we get the entailment from ``John wrote a
book'' to ``John was writing a book''.\footnote{Note that this is a different interpretation of the usual accounts of the progressive in the literature which take a progressive sentence  to be  true at an interval $t_0$
 iff there is an interval $t_1$ such that $t_0$ is a non final
subinterval of $t_1$ and the progressive sentence is true at $t_1$ \citep{ogihara2007tense}. In terms of implementation, the solution we provide here seems to be simpler.}

Finally we note that not all verbs are subject to the
stative/achievement distinction induced by the progressive. For
example, the phrases ``John ran'' and ``John was running'' appear to be
logically equivalent, for entailment purposes.

@@

Another thing we are willing to try is an extension of the FraCaS
test suite that will include more fine-grained examples of lexical and
grammatical aspect, as well as the interaction between those two, for example
cases where one needs to actually distinguish between achievements and
accomplishments on the basis of their inferential properties:

@@

Further cases to be included in an extended FraCaS future suite
involve examples where the interaction between different tenses needs
to be captured:
%Another issue to be looking at is the codification of lexical
%aspect.\todo{update} In what we have been doing so far, we did not distinguish
%between the present continuous and the present or the simple past and
%the past continuous and their interaction with non-lexical aspect.
%The main reason for this is that in the test suite, even though there
%are examples of inferences based on the either lexical or non-lexical
%aspect, no example where an interaction between the two exists. For
%example, one such case is the one shown below and exemplifies the
%difference between activities and accomplishments with respect to the
%different inferences they give when used in the past continuous:

@@

% Things we cannot handle (I think): John was writing a book in 1991
% → John wrote a book in 1991 (I think we predict entailment
% here). This is because there is no distinction with progressive and
% non progressive aspect. We did not do it, I guess because it was
% needed, but just for future reference. There are similar examples
% playing around with LEXICAL aspect (aspect grammatically encoded)
% that we could look at in future work

% Accomplishments usually involve a culmination point in the action
% which is necessary for the event to be completed. This is not
% necessary for activities. John stopped writing a book not→ John
% wrote a book (but does imply that John was writing a book), but,
% John stopped running → John run and John was runnning

>>>>

We treat verbs in the progressive form as different semantically from
the non-progressive form. For example, ``John was writing a book'' is
encoded as
$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, PROG\_write(John,book,t_1,t_2)$,
while ``John wrote a book'' is encoded as
$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, write(John,book,t_1,t_2)$. This distinction is
necessary because in our analysis the progressive form ($PROG\_write$) is subject to
$\constant{subsumption}$. That is, if John is writing in the interval
$[t_1,t_2]$ then he is writing in any sub-interval of
$[t_1,t_2]$. This interpretation corresponds to the idea the the action takes place continuously over the whole interval.
However, the same cannot be said of the non-continuous form ($write$): the
end-points of the interval indicate the time needed to complete the
achievement. (For example, ``John wrote a book in 1993'' neither
entails ``John wrote a book in January 1993'' nor ``John wrote a book
in December 1993''.) (In fact, \emph{write}, in the non-progressive
from, is on the contrary subject to $\constant{unicity}$.) Finally, we
also have $write(x,y,t_1,t_2) → PROG\_write(x,y,t_1,t_2)$. That is,
the achievement (or \emph{activity} in our terminology) variant implies the stative variant, for the
same interval.  Consequently we get the entailment from ``John wrote a
book in 1993'' to ``John was writing a book in 1993'', but not the other way around.

We note however that this interpretation differs only slightly from
the usual accounts of the progressive in the
literature. \Citet{ogihara2007tense} summarizes the position of
\citet{bennett1978toward} as follows: a progressive sentence is true
at an interval $[t_0,t_1]$ iff there is an interval $[t'_0,t'_1]$ such
that $[t_0,t_1]$ is a non-final subinterval of $[t'_0,t'_1]$ and the
progressive sentence is true at $[t'_0,t'_1]$. This is very similar to
our approach ($\constant{subsumption}$ for the progressive form only),
but there is a difference regarding final intervals. Yet in our view
this difference is hard to justify: we cannot see why ``John was
writing a book in 1993'' entails that he was writing it January,
February, etc. but not in December.

\Citet{ogihara2007tense} argues that this simple account of the
progressive fails to reject a sentence such as ``Lee is resembling
Terri.'' while ``Lee is walking'' is acceptable. We argue instead that
the latter should be rejected for pragmatic reasons. Indeed, when a
predicate holds for a very long interval, one typically uses the
simple present tense in English. Therefore the continuous form
pragmatically implies that the predicate holds for a limited interval.
But, without further context, the predicate ``resemble Terri'' does
not vary over time (while ``walk'' generally does). Therefore the
continuous form ``Lee is resembling Terri'' is confusing: one implies
a limited interval, but the semantics of resembling normally yield an unlimited interval.
%
Because we do not account for pragmatics, we prefer to retain the
simplest account based on $\constant{subsumption}$.

Finally we stress that not all verbs are subject to the
stative/achievement distinction induced by the progressive. For
example, the phrases ``John ran'' and ``John was running'' appear to
be logically equivalent, for entailment purposes.

@@

Another thing that we are willing to try is an extension of the FraCaS
test suite that will include more fine-grained examples of lexical and
grammatical aspect, as well as the interaction between those two, for example
cases where one needs to actually distinguish between achievements and
accomplishments on the basis of their inferential properties:

@@

Further cases to be included in an extended FraCaS future suite
involve examples where the interaction between different tenses needs
to be captured:

%Another issue to be looking at is the codification of lexical
%aspect.\todo{update} In what we have been doing so far, we did not distinguish
%between the present continuous and the present or the simple past and
%the past continuous and their interaction with non-lexical aspect.
%The main reason for this is that in the test suite, even though there
%are examples of inferences based on the either lexical or non-lexical
%aspect, no example where an interaction between the two exists. For
%example, one such case is the one shown below and exemplifies the
%difference between activities and accomplishments with respect to the
%different inferences they give when used in the past continuous:

@@
